---
title: "Take-home_Ex02"
author: "Zou Jiaxun"
date: "Oct 2, 2024"
data-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
  cache: true
format:
  html:
    code-fold: true
    code-summary: "Click to view code"
---

# **1 Overview**

## **1.1 Background**

Tourism is a critical pillar of Thailand’s economy, contributing around 20% of the country’s GDP. In 2019, the sector generated approximately 90 billion US dollars from both domestic and international tourism. However, the COVID-19 pandemic severely impacted the industry, reducing revenues to 24 billion US dollars in 2020. Since September 2021, tourism revenue has shown a steady recovery, as indicated by the data from January 2019 to February 2023. Despite this rebound, the benefits of the tourism economy are concentrated in a few key provinces—Bangkok, Phuket, Chiang Mai, Sukhothai, and Phetchaburi—highlighting an uneven distribution of tourism income across the country.

## **1.2 Objectives**

::: {.panel-tabset}

## Aim

Test **tourism economy** of Thailand are **independent** from space and space and time or not.

Detect where are the **clusters and outliers**, and the emerging **hot spot/cold spot areas**.

## Test

Using appropriate function of **sf** and **tidyverse**, preparing the following geospatial data layer:

-   a study area layer in sf polygon features. It must be at [province level](http://localhost:3980/Take-Home_Ex/Take-home_Ex02/Take-home_Ex02.html#0) (including Bangkok) of Thailand.

-   a tourism economy indicators layer within the study area in sf polygon features.

-   a derived tourism economy indicator layer in [**spacetime s3 class of sfdep**](http://localhost:3980/Take-Home_Ex/Take-home_Ex02/Take-home_Ex02.html#0). Keep the time series at **month and year levels**.

Using the extracted data, perform

-   global spatial autocorrelation analysis

-   local spatial autocorrelation analysis

-   emerging hotspot analysis

by using [sfdep methods](http://localhost:3980/Take-Home_Ex/Take-home_Ex02/Take-home_Ex02.html#0).

Describe the spatial patterns revealed by the analysis above.

:::

# **2 Data Wrangling**

## **2.1 Loading Packages**

The following packages will be used for this exercise:

```{r}
pacman::p_load(sf, spNetwork, tmap, tidyverse, dplyr,
               spatstat, knitr, sfdep, spdep, ClustGeo,
               ggpubr, cluster, factoextra, NbClust,
               heatmaply, corrplot, psych, GGally)

package_info <- data.frame(
  Package = c(
    "sf", "spNetwork", "tmap", "tidyverse", "dplyr", 
    "spatstat", "knitr", "sfdep", "spdep", "ClustGeo", 
    "ggpubr", "cluster", "factoextra", "NbClust", 
    "heatmaply", "corrplot", "psych", "GGally"
  ),
  Description = c(
    "For importing, managing, and handling geospatial data",
    "For network-based spatial analysis, including kernel density estimation on networks",
    "For thematic mapping and interactive visualizations",
    "For non-spatial data wrangling and general data analysis with a collection of packages",
    "For data manipulation and pipeline tools, part of the tidyverse",
    "For spatial statistics and point pattern analysis",
    "For dynamic report generation using LaTeX, HTML, and Markdown",
    "For spatial statistics with a focus on spatial dependence and spatial autoregressive models",
    "For spatial dependence analysis in spatial econometrics, including spatial lag and error models",
    "For clustering spatial data using hierarchical clustering and geographical constraints",
    "For creating publication-ready plots",
    "For performing various clustering algorithms and analysis",
    "For visualizing and interpreting the results of multivariate data clustering",
    "For determining the optimal number of clusters in a dataset using various indices",
    "For creating interactive heatmaps and hierarchical clustering visualizations",
    "For visualizing correlation matrices using various methods",
    "For performing descriptive statistics, factor analysis, and psychometrics",
    "For visualizing correlations, pair plots, and creating enhanced data analysis plots"
  )
)

knitr::kable(package_info, col.names = c("Package", "Description"))
```

**The Data**

The following data are used for this study:

-   Aspatial:

    -   *Thailand Domestic Tourism Statistics* at Kaggle. You are required to use version 2 of the data set.

-   Geospatial

    -   *Bus Stop Location* from [LTA DataMall](https://datamall.lta.gov.sg/content/datamall/en/static-data.html). It provides information about all the bus stops currently being serviced by buses, including the bus stop code (identifier) and location coordinates.

    -   *hexagon*, a [hexagon](https://desktop.arcgis.com/en/arcmap/latest/tools/spatial-statistics-toolbox/h-whyhexagons.htm) layer of 250m to replace the relative coarse and irregular Master Plan 2019 Planning Sub-zone GIS data set of URA.

```{r}
tourism <- read_csv("data/aspatial/thailand_domestic_tourism_2019_2023_ver2.csv")
bd <- st_read(dsn = "data/geospatial", layer = "tha_admbnda_adm2_rtsd_20220121")
st_crs(bd)

bd <- st_transform(bd, crs = 32647)
st_crs(bd)

any(duplicated(bd))

bd_clean <- bd %>%
  select(where(~ !all(is.na(.))))
```

# **3 Geospatial Analysis**

## **3.1 if the key indicators of tourism economy of Thailand are independent from space and space and time.**

# **4 Geovisualisation & Geocommunication**

# **5 Reproducibility**

# **6 Bonus**

# **7 Reference**
