---
title: "1  Geospatial Data Science with R"
---

Geospatial Data Science is a process of importing, wrangling, integrating, and processing geographically referenced data sets. In this hands-on exercise, we will learn how to perform geospatial data science tasks in R by using **sf** package.

By the end of this hands-on exercise, we should acquire the following competencies:

-   installing and loading [sf](https://r-spatial.github.io/sf/) and [tidyverse](https://www.tidyverse.org/) packages into R environment,

-   importing geospatial data by using appropriate functions of **sf** package,

-   importing aspatial data by using appropriate function of **readr** package,

-   exploring the content of simple feature data frame by using appropriate **Base R** and **sf** functions,

-   assigning or transforming coordinate systems by using appropriate *sf* functions,

-   converting an aspatial data into a sf data frame by using appropriate function of **sf** package,

-   performing geoprocessing tasks by using appropriate functions of **sf** package,

-   performing data wrangling tasks by using appropriate functions of **dplyr** package and

-   performing Exploratory Data Analysis (EDA) by using appropriate functions from **ggplot2** package.

> Note: we are encouraged to read the reference guide of each function, especially the input data requirements, syntaxt and argument option before using them.

## **1.2 Data Acquisition**

Data are key to data analytics including geospatial analytics. Hence, before analysing, we need to assemble the necessary data. In this hands-on exercise, we are required to extract the necessary data sets from the following sources:

-   Master Plan 2014 Subzone Boundary (Web) from [data.gov.sg](https://data.gov.sg/)

-   Pre-Schools Location from [data.gov.sg](https://data.gov.sg/)

-   Cycling Path from [LTADataMall](https://www.mytransport.sg/content/mytransport/home/dataMall.html)

-   Latest version of Singapore Airbnb listing data from [Inside Airbnb](http://insideairbnb.com/get-the-data.html)

> Note: The purpose of this section is not merely extracting the necessary data sets. It also aims to introduce we to public available data sets. We are encouraged to explore the rest of the available data sets in these three data sources.

### **1.2.1 Extracting the geospatial data sets**

Next, at the `Hands-on_Ex01` folder, create a sub-folder called `data`. Then, inside the `data` sub-folder, create two sub-folders and name them `geospatial` and `aspatial` respectively.

Place `Master Plan 2014 Subzone Boundary (Web)`, `Pre-Schools Location` and `Cycling Path` zipped files into `geospatial` sub-folder and unzipped them. Copy the unzipped files from their respective sub-folders and place them inside `geospatial` sub-folder.

### **1.2.2 Extracting the aspatial data set**

Now, we will extract the downloaded `listing` data file. At `Downloads`folder, cut and paste `listing.csv` into `aspatial` sub-folder.

## **1.3 Getting Started**

In this hands-on exercise, two R packages will be used. They are:

-   **sf** for importing, managing, and processing geospatial data, and

-   **tidyverse** for performing data science tasks such as importing, wrangling and visualising data.

Tidyverse consists of a family of R packages. In this hands-on exercise, the following packages will be used:

-   **readr** for importing csv data,

-   **readxl** for importing Excel worksheet,

-   **tidyr** for manipulating data,

-   **dplyr** for transforming data, and

-   **ggplot2** for visualising data

Type the following code chunk.

```{r}
pacman::p_load(sf, tidyverse)
```

What we can learn from the code chunk above:

-   [`p_load`](https://cran.r-project.org/web/packages/pacman/pacman.pdf) function pf [**pacman**](https://github.com/trinker/pacman) package is used to install and load sf and tidyverse pacages into R environment.

## **1.4 Importing Geospatial Data**

In this section, we will learn how to import the following geospatial data into R by using [*st_read()*](https://r-spatial.github.io/sf/reference/st_read.html) of **sf** package:

-   `MP14_SUBZONE_WEB_PL`, a polygon feature layer in ESRI shapefile format,

-   `CyclingPath`, a line feature layer in ESRI shapefile format, and

-   `PreSchool`, a point feature layer in kml file format.

### **1.4.1 Importing polygon feature data in shapefile format**

The code chunk below uses *st_read()* function of **sf** package to import `MP14_SUBZONE_WEB_PL` shapefile into R as a polygon feature data frame. Note that when the input geospatial data is in shapefile format, two arguments will be used, namely: `dsn` to define the data path and `layer` to provide the shapefile name. Also note that no extension such as .shp, .dbf, .prj and .shx are needed.

```{r}
mpsz = st_read(dsn = 'data/geospatial', layer = 'MP14_SUBZONE_WEB_PL')
```

The message above reveals that the geospatial objects are multipolygon features. There are a total of 323 multipolygon features and 15 fields in `mpsz` simple feature data frame. `mpsz` is in **svy21** projected coordinates systems. The bounding box provides the x extend and y extend of the data.

### **1.4.2 Importing polyline feature data in shapefile form**

The code chunk below uses *st_read()* function of **sf** package to import `CyclingPath` shapefile into R as line feature data frame.

```{r}
cyclingpath = st_read(dsn = 'data/geospatial', layer = 'CyclingPathGazette')
```

The message above reveals that there are a total of 2558 features and 2 fields in `cyclingpath` linestring feature data frame and it is in **svy21**projected coordinates system too.

### **1.4.3 Importing GIS data in kml format**

The `PreSchoolsLocation` is in kml format. The code chunk below will be used to import the kml into R. Notice that in the code chunk below, the complete path and the kml file extension were provided.

```{r}
preschool = st_read("data/geospatial/PreSchoolsLocation.kml")
```

The message above reveals that `preschool` is a point feature data frame. There are a total of 2290 features and 2 fields. Different from the previous two simple feature data frame, preschool is in **wgs84** coordinates system.

## **1.5 Checking the Content of A Simple Feature Data Frame**

In this sub-section, we will learn different ways to retrieve information related to the content of a simple feature data frame.

### **1.5.1 Working with *st_geometry()***

The column in the sf data.frame that contains the geometries is a list, of class `sfc`. We can retrieve the geometry list-column in this case by mpsz\$geom or mpsz\[\[1\]\], but the more general way uses *st_geometry()* as shown in the code chunk below.

```{r}
st_geometry(mpsz)
```

Notice that the print only displays basic information of the feature class such as type of geometry, the geographic extent of the features and the coordinate system of the data.

### **1.5.2 Working with glimpse()**

Beside the basic feature information, we also would like to learn more about the associated attribute information in the data frame. This is the time you will find *glimpse()* of **dplyr**. very handy as shown in the code chunk below.

```{r}
glimpse(mpsz)
```

*glimpse()* report reveals the data type of each fields. For example `FMEL-UPD_D` field is in **date** data type and `X_ADDR`, `Y_ADDR`, `SHAPE_L` and `SHAPE_AREA` fields are all in **double-precision values**.

### **1.5.3 Working with *head()***

Sometimes we would like to reveal complete information of a feature object, this is the job of [*head()*](https://www.rdocumentation.org/packages/utils/versions/3.6.2/topics/head) of Base R

```{r}
head(mpsz, 5)
```

> Note: One of the useful argument of *head()* is it allows user to select the numbers of record to display (i.e. the n argument).

## **1.6 Plotting the Geospatial Data**

In geospatial data science, by looking at the feature information is not enough. We are also interested to visualise the geospatial features. This is the time will find *plot()* of R Graphic comes in very handy as shown in the code chunk below.

```{r}
plot(mpsz)
```

The default plot of an sf object is a multi-plot of all attributes, up to a reasonable maximum as shown above. We can, however, choose to plot only the geometry by using the code chunk below.

```{r}
plot(st_geometry(mpsz))
```
